---
title: "Association Rules "
author: " "
date: "July 21, 2017"
output: html_document
---
# Problem 3 - Assignment
### Association Rules for "Flight Delay" dataset: Generate the rules and identify the patterns. 

```{r}
rm(list = ls(all = TRUE))
```

1. Read the data into R

```{r, echo=T}
setwd("I:/DATA-SCIENCE/Insofe/Assignments/AssociationRules")
flight.data = read.csv("FlightDelays.csv", header = TRUE)
```


2. Look at the summary of all the variables and convert the following variables as factors 
        + Weather 
        + DAY_WEEK 
        + Flight Status
```{r}
summary(flight.data)
str(flight.data$CRS_DEP_TIME)

flight.data$Weather = as.factor(as.character(flight.data$Weather))
flight.data$DAY_WEEK = as.factor(as.character(flight.data$DAY_WEEK))
flight.data$Flight.Status = as.factor(as.character(flight.data$Flight.Status))

str(flight.data)
```

3. Bin the numeric variable 'CRS_DEP_TIME' into 4 bins as follows: If time is between 6 AM to 12 noon code as 1 and If the time is between 2 , etc.
```{r}
flight.data$CRS_DEP_TIME = cut(flight.data$CRS_DEP_TIME, c(1159,1759,2359,559), labels = c(1,2,3))
summary(flight.data)

```

4. Convert the data frame in a transactions object. Look at the first 6 transactions to understand how 'arules' library requires data to be transformed. Use the below R code to view the transactions data 
```{r}
library(arules)
flight.transaction = as(flight.data, "transactions")

image(flight.transaction)
itemFrequency(flight.transaction)
```

5. Apply 'arules' algorithm and play with various support, lift and confidence values.
```{r}
rulesA<-apriori(flight.transaction, parameter = list(support = 0.1, conf = 0.2, target = 'rules', minlen = 3))
```

6. Inspect all the rules.
```{r}
inspect(rulesA[1:10])
```

7. Filter the rules with specific LHS and RHS conditions 
E.g.; Filter the rules with Flighstatus=0
```{r}
rules.filterA = sort(subset(rulesA, subset = rhs %in% 
                        c("Flight.Status=0", "Flight.Status=1")), by=c("support", "confidence"))
rules.sorted = sort(rulesA, by="lift")
inspect(rules.sorted)
```

8. Filter redundant rules if any
```{r}
subset.matrix <- is.subset(rules.sorted, rules.sorted,sparse = FALSE)
subset.matrix[lower.tri(subset.matrix, diag=T)] <- NA
redundant <- colSums(subset.matrix, na.rm=T) >= 1
which(redundant)

rules.pruned <- rules.sorted[!redundant]
inspect(rules.pruned)
```

9. Pot and visualize the rules
```{r}
library(arulesViz)
plot(rules.pruned)

```